---
title: "Effects of land use and land cover changes on biodiversity, ecosystem services, and carbon storage in Norway: A systematic map and decision support tool"
format:
  html:
    toc: true
params:
  version: "0.1"
  search-results: 36418
editor: visual
---

## 

Version: `r params$version`

# Background

Land Use and Land Cover Change (LULCC) is considered the dominant driver of recent biodiversity loss both globally and in Norway, and directly affects nature’s ability to provide natural goods and services (Díaz et al., 2019; Malhi et al., 2020). LULCC also affects an ecosystem’s ability to sequester and store carbon in the soil and biological material, and therefore directly affects an ecosystem’s ability to assist in the mitigation of, and adaptation to, climate change. Quantifying, understanding, and communicating the dynamics of land change and its effects on biodiversity, ecosystem services, and carbon sequestration is therefore a premise for evidence-based land use policies. However, we currently lack systematic evidence on the topic from Norwegian ecosystems. Much of the public discourse at the national level is dominated by international global knowledge summaries, despite significant uncertainty around downscaling and transferring knowledge from global and regional syntheses to the local/national level, across ecosystems and taxonomic groups. Previous knowledge syntheses based on data from Norwegian ecosystems has either been following specific approaches (e.g. Ecosystem Assessment; (Nybø et al., 2017)), largely based on expert judgements (Artsdatabanken, 2021), or limited in scope. The lack of a comprehensive overview of research-based knowledge from Norwegian ecosystems also leads to different interest groups and sectors relying on different "facts," often leading to escalating conflicts and poor decisions. Therefore, there is a pressing need both for a summary of knowledge and the identification of real knowledge gaps.

Politics and decision making related to land use and its effects on biodiversity and climate are currently among the most conflict-ridden and hotly debated political issues in Norway. The conflicts typically involve dilemmas between different societal concerns and interests supported by different sectors and actor groups (Guldbrandsen & Handberg, 2022; Skogen et al., 2021).  Knowledge controversy is also often a central trait of the conflicts, involving disagreement on the status of knowledge and conflicting knowledge claims about the environmental effects of different options of land use (Aspøy & Stokland, 2022; Rusch et al., 2022). IPBES (2019, 2022a, 2022b) and IPCC (2018) have proposed governance approaches and measures, such as those related to transformative change and transformative governance, to enable the achievement of the SDGs, the 1.5℃ limit, and the 2050 Vision for Biodiversity. Such approaches involve improved integration and interaction between societal goals, governance sectors, and multiple knowledge sources, and are thus highly relevant for Norwegian politics and decision making related to land use. However, the utilization and adaption of internationally recommended approaches and tools in a Norwegian context is currently hindered by the lack of a comprehensive understanding of national and local land use-related dilemmas and conflicts, as well as currently employed tools and approaches. Therefore, there is an urgent need for a knowledge synthesis and identification of knowledge gaps concerning these aspects of land use and its effects on biodiversity and climate.

Systematic maps (also known as Evidence gap maps or Evidence maps) were designed to give an overview of the available evidence on a broad topic (Saran et al., 2020). They do not seek to answer narrow well-defined questions (as systematic reviews do) but rather to provide a means to identify and catalogue available evidence on policy-relevant questions (James et al., 2016). Systematic maps have been realised in a variety of different ways but generally provide the results in a user-friendly fashion such as graphs, maps or searchable databases (Miake-Lye et al., 2016). Increasingly interactive plotting and mapping is being used to provide a dynamic output to users (such as EviAtlas; Haddaway et al., 2019). It is important that systematic maps are not portrayed as an end in themselves. Without further evidence synthesis decision makers need to be cautious about making decisions based on the weight of evidence alone (i.e.  vote-counting, c.f (Hedges & Olkin, 1980). Systematic maps cannot replace the rigor of systematic reviews but can give decision-makers an understanding of the evidence base by highlighting gaps in, or clusters of, evidence. They can highlight where evidence synthesis can take place (i.e., there is enough available evidence for a systematic review) or where primary research is required (i.e. there is a lack of evidence). Decision models can supplement the results of a systematic evidence synthesis to place the available evidence into the decision context (Carrick et al., 2019).

## Stakeholder engagement

The outline for the review was presented at a conference held by the Norwegian Environment Agency and the Norwegian Research Council October 23, 2022. A general stakeholder meeting was arranged by NINA on April 25, 2023. The focus and scope of the review was narrowed and better defined during a specific stakeholder event on #date. Details of this meeting and the modifications in scope are available on request. Stakeholders who attended this event were invited to comment on the draft protocol prior to submission for publication. Stakeholders were not engaged during the conduct of the review.

## Objective of the map

The main goal of this project is to provide a comprehensive overview of available knowledge related to the effects of land use and land use changes on Norwegian nature, with a focus on biodiversity, ecosystem services, climate adaptation, carbon sequestration and storage. The map is restricted in geographical scope to Norway, Sweden and Finland (see inclusion criteria below). Moreover, the report is accompanied by a decision support tool as an extension of the systematic map to contextualize the available knowledge. The tool will be designed as a conceptual model of the field and highlight the importance of knowledge gaps in different areas.

### Primary question

The main research question in our systematic map is ‘What evidence exist about the impacts of land use and land cover change on biodiversity, ecosystem services and carbon sequestration in Norway, what type of conflicts have been identified between different forms of land use, and what tools are suggested to improve governance".

### Secondary questions

In addition, we will investigate what characterizes the existing body of literature on the topic, in terms of:

-   What proportion of the research is based on empirical data 

-   What proportion of the research is based on experimental vs observational 

-   What proportion of the research is based on qualitative vs quantitative/mixed 

-   How is research spread across the major ecosystem types/realms 

-   How is research spread across the disciplines 

-   How is research spread across level of biological organization (species, taxa, ecosystem or services) 

-   What are the dominant types of land use that are investigated? 

-   What regions (in Norway) are most represented 

-   What proportion of the research focuses on governance, conflicts and tools?  

-   What are the dominant/marginal social scientific analytical approaches of the studies? 

-   What type of actors are most represented in the studies? 

-   Which sectors and societal concerns are most studied? 

-   Which level of societal scale do the studies represent?  

-   Which land use conflicts are most researched? 

-   Which tools are proposed and investigated? 

# Methods

## Deviations from protocol

## Search for literature

### Online publication databases

Systematic literature searches for academic literature were conducted in Scopus, and in the following Web of Science Core Collection databases:

-   Science Citation Index Expanded (SCI-EXPANDED): 1987–present.

-   Social Sciences Citation Index (SSCI): 1987–present.

-   Arts and Humanities Citation Index (AHCI): 1987–present.

-   Emerging Sources Citation Index (ESCI): 2015–present.

Several preliminary searches with different search terms were performed in Web of Science, which led to adjustments in the search string (See [the protocol](https://osf.io/4ad3x)).

### Search terms and string

We used the PICO framework (James et al., 2016) to structure our search with the search terms and categories derived from the primary research question.

The following search string was used in WoS and a similar string used in Scopus (adjusted to the interface):

<!--# Add link to search strings -->

*TS=("land use" OR "land cover" OR "space use" OR LULCC OR (planning AND (spati\* OR region\* OR area\* OR conservation OR land\* OR urban)) OR (conversion AND (land\* OR habitat\* OR area\* OR ecosystem\* OR natur)) OR (change NEAR/3 (land\* OR area)) OR "national park" OR landscape\* OR habitat\* OR Husbandry OR "land shar" OR "land spar" OR (Outdoor NEAR/3 (life OR livin\* OR recreat)) OR (restoration AND (nature OR ecosystem)) OR forestry OR (reindeer NEAR/3 (herd\* OR husbandr)) OR (livestock NEAR/3 graz) OR "holiday home" OR "secondary home"OR (recreation\* NEAR/3 (home\* OR cottage\* OR cabin\* OR propert)) OR tourism OR "wind power" OR "Land manag" OR "Nature conservation" OR ((governance OR management) AND (spati\* OR region\* OR area\* OR conservation OR land\* OR urban)) OR aquaculture OR "fish farm" OR hydropower OR “green energy” OR “energy development” OR herding OR agriculture OR transportation OR road OR “oil and gas” OR petroleum OR mining OR windpower OR traffic OR road\* OR railway\* OR “hydroelectric power” OR “solar power” OR “solar electricity” OR “power plant” OR LULC OR LULUCF OR AFOLU OR FOLU)*

*AND*

*TS=(sustainab\* OR biodiversity OR "Ecosystem service" OR (carbon NEAR/3 (stor OR sequest)) OR (ecosystem NEAR/3 (impact\* OR effect\* OR provision\* OR regulat\* OR support\* OR cultur)) OR ((loss OR shrink OR bisection\* OR perforation OR fragmentation) AND (habitat OR land)) OR (climate NEAR/3 adapt) OR ecodiversity OR "ecosystem type" OR "ecological process" OR anthropogenic OR conflict OR (diversity NEAR/3 (ecological OR speci\* OR population\*)) OR red-list OR "water purification" OR "soil retention" OR spiritual OR cultural OR aesthetics OR "climate mitigation")*

We screened all systematic reviews and systematic maps from the international literature, the two categories were combined with TS=("Systematic review*" OR "systematic map*" OR "rapid review\*"). For the primary literature, the two categories were combined with TS=(Norway OR Norwegian OR Sweden OR Swedish OR Finland OR Finnish), and all results screened.

### Languages

Studies written in English, Norwegian, Swedish or Danish were included in the review. Studies in other languages were not included in the review.

### Internet searches

### Additional searches

Articles cited by and that cite (forward and backward citation chasing) the 15 benchmark articles were searched for using the citationchaser Shiny App (Haddaway et al. 2022).

### Testing the comprehensiveness of the search

Of the 15 benchmarks, 2 were not recovered in the searches.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Overlap between the different sources of literature and the benchmark papers"

library(CiteSource)
library(tidyverse)
unique_citations <- readRDS("C:/Users/matthew.grainger/Documents/Projects_in_development/landuse_effects_biodiversity/data/unique_citations.RDS")
source_comparison <- compare_sources(unique_citations, comp_type = "sources")
# Generate a source comparison upset plot.
plot_source_overlap_upset(source_comparison, decreasing = c(TRUE, TRUE))
```

### Assembling and managing search results

### Article screening

We used pre-defined inclusion/exclusion criteria to screen eligible studies. Included studies were de-duplicated based on DOI and title matches using the R package 'ASySD' (Hair et al. 2021). We retained the latest version of any version-controlled articles (e.g. published versions of pre-prints were retained).  

After de-duplication, title and abstract screening was done using the eligibility criteria, and any studies classified as uncertain were assessed at the full-text stage. The screening was done using Rayyan, with double-screening by a primary and several secondary reviewers. After 50 include/exclude decisions we used the Rayyan-AI 5 star rating to rank the abstracts for assessment. This algorithm assesses abstracts and titles using natural language processing to identify records that have a high potential for inclusion (5 star) or those with a lower potential (1 star).

Weekly monitoring was done for group disagreement rates, and the eligibility criteria were re-evaluated and readjusted when the rate exceeded 15%. Conflicting decisions were reconciled through discussions with the whole review group. Studies tagged as unsure at each stage were assessed at the next level, with those remaining tagged as unsure reviewed by the entire team.  

### Eligibility criteria

We reviewed the collated studies obtained from the searches using the criteria set below. Only studies conducted in Norway, Sweden and Finland or systematic reviews (also referred to as meta-analyses) with global, European or Fennoscandian coverage were included in the review. Only records where the full-text is written in English, Norwegian, Swedish, or Danish were considered eligible.

Studies will be eligible if they assess land use or land-use-change effects on the biodiversity, ecosystem services/functions or carbon sequestration and storage in Feno-Scandinavia. Only studies that focus on terrestrial, aquatic or coastal areas are include (i.e. studies focusing on marine systems are excluded).  

Studies will also be eligible if they assess conflicts, tools or governance relating to land use and cover change its effects on biodiversity, ecosystem services and/or carbon sequestration and storage by ecosystems in Fenno-Scandinavia. The relation should be explicit in study for the record to be included.  

Studies of all scales (national, regional, and local) will be considered eligible.  

2.  **Eligible interventions**

Land use and/or land cover change through one or multiple drivers (see Appendix) which are anthropogenic in origin and result in one or more of the core mechanisms (change in habitat – amount; habitat composition; habitat structure; habitat – function; indirect, e.g. changed human use in recreation etc). 

3.  **Eligible outcomes**

Eligible outcomes will include function, structure, diversity, abundance, behaviour, composition, or richness in any level of biological organisation (genetic, species, community, ecosystem, landscape – multiple ecosystems), as well as studies focusing on the amount of, diversity or category of ecosystem services (provisioning, regulating, supporting or cultural service and, carbon storage and sequestration).    

In addition, studies within the social sciences will be included if they explicitly address conflicts, tools and governance pertaining to effects land use and land cover change on biodiversity, ecosystem services and/or carbon sequestration and storage.    

4.  **Eligible types of study design**

All study designs will be considered eligible, including both quantitative and qualitative studies.  

### Study validity assessment

### Data coding strategy

### Data mapping method

# Review findings

## Results of the search for literature

```{r, Prisma statement}
#| echo: false
#| message: false
#| warning: false
library(PRISMAstatement)

prisma(
  found = params$`search-results`,
  found_other = 1965,
  no_dupes = 20265,
  screened = 14808, # Needs updating
  screen_exclusions = 12144,
  full_text = 2398,
  full_text_exclusions = 0,
  qualitative = 0,
  quantitative = 0,
  width = 800, height = 800
)
```

### Sources and document types of included articles

## Review descriptive statistics

## Norway papers

```{r}
#| echo: false
#| message: false
#| warning: false
# Extract data and plot from the extraction sheet ----
## load libraies----
library(tidyverse, quietly = TRUE)
library(googlesheets4, quietly = TRUE)
library(rayyanR)
library(synthesisr)
library(openalex)
library(openalexR)
# gs4_auth()
## read in GoogleSheets----

# form<-read_sheet("https://docs.google.com/spreadsheets/d/1lhZtZ5oSfTek7iaWWv8r40muWBEAJ76WsGzOXcb9Da0")# copy of sheet to test

# form <- read_sheet("https://docs.google.com/spreadsheets/d/1wwOfwz_QqwXpNxKx4Jupn336x2GttOfoXKibd-_rMuc")
# 
# ## clean names----
# form <- form |>
#   janitor::clean_names()
# ## explore----
# names(form)
# 
# ### select Norwegian studies----
# Nor_form <- form |>
#   filter(grepl("Norway", country))
# 
# ### selet studies that are included in the review----
# Nor_form_inc <- Nor_form |>
#   filter(does_the_study_assess_land_use_or_land_use_change_effects_on_the_biodiversity_ecosystem_services_functions_or_carbon_sequestration_and_storage_in_terrestrial_aquatic_or_coastal_areas_are_include_i_e_studies_focusing_on_marine_systems_are_excluded == "Yes" | does_the_study_assess_conflicts_tools_or_governance_relating_to_land_use_and_cover_change_its_effects_on_biodiversity_ecosystem_services_and_or_carbon_sequestration_and_storage_by_ecosystems_in_fenno_scandinavia_the_relation_should_be_explicit_in_study_for_the_record_to_be_included == "Yes")
# 
# ### Get matched details from Rayyan----
# 
# # read the Rayyan results
# 
# rayyan_biblio <- synthesisr::read_ref("data/Extraction/articles.ris")
# rayyan_biblio_cleaned <- rayyanR::parse_rayyan(rayyan_df = rayyan_biblio)
# 
# Nor_form_inc$accession_zr<-unlist(Nor_form_inc$study_id)
# 
# join_Nor_form <- Nor_form_inc |>
#   left_join(rayyan_biblio_cleaned, join_by("accession_zr"))
# 
# # join_Nor_form |> view()
# 
# join_Nor_form |>
#   filter(is.na(join_Nor_form$doi)) |>
#   select(study_id, study_title)
# 
# #### find missing doi's by title
# which(is.na(join_Nor_form$doi))
# #11, 35, 47,52,60,61
# 
# works_search <- oa_fetch(
#   entity = "works",
#   title.search = c("Roads forestry wolves"),
#   verbose = TRUE
# )
# works_search$doi
# # "https://doi.org/10.1002/ecs2.3358"
# 
# 
# works_search <- oa_fetch(
#   entity = "works",
#   title.search = c("CONFLICT AND COMPLEXITY - THE CASE OF HYDROPOWER CONSTRUCTION IN NORWAY"),
#   verbose = TRUE
# )
# works_search$doi
# # "https://doi.org/10.3233/hsm-1986-6105"
# works_search <- oa_fetch(
#   entity = "works",
#   title.search = c("Influence of site and stand factors on the distribution of crustose lichens of the Caliciales in a suboceanic spruce forest area in central Norway"),
#   verbose = TRUE
# )
# works_search |> view()
# works_search$doi
# # "https://doi.org/10.1017/s0024282996000424"
# 
# works_search <- oa_fetch(
#   entity = "works",
#   title.search = c("Lichens in spruce forest stands of different successional stages in central Norway with emphasis on diversity and old growth species"),
#   verbose = TRUE
# )
# 
# works_search$doi
# works_search <- oa_fetch(
#   entity = "works",
#   title.search = c("Impacts of experimental habitat fragmentation on ground beetles"),
#   verbose = TRUE
# )
# 
# works_search
# 
# 
# works_search <- oa_fetch(
#   entity = "works",
#   title.search = c("Cumulative impacts of tourist resorts on wild reindeer"),
#   verbose = TRUE
# )
# 
# works_search$doi
# 
# join_Nor_form[11, ]$doi <- "10.3233/hsm-1986-6105"
# join_Nor_form[35, ]$doi <- "10.1002/ecs2.3358"
# join_Nor_form[47, ]$doi <- "10.1017/s0024282996000424"
# join_Nor_form[52, ]$doi <- "10.1127/nova.hedwigia/66/1998/283"
# join_Nor_form[60, ]$doi <- NA
# join_Nor_form[61, ]$doi <- "10.14430/arctic829"
# 
# # get openAlexR data
# 
# 
# doi_list <- strsplit(join_Nor_form$doi, split = " ")
# doi_list <- sapply(doi_list, "[[", 1)
# 
# # oa_fetch(
# #   doi = c(doi_list),
# #   entity = "works",
# #   verbose = TRUE
# # ) %>%
# #   show_works() %>%
# #   knitr::kable()
# 
# bib_ls <- list(
#   doi = c(doi_list),
#   entity = "works"
# )
# 
# # oa_fetch(
# #   # identifier = c("W2741809807", "W2755950973"),
# #   doi = c("10.1016/j.joi.2017.08.007", "https://doi.org/10.1093/bioinformatics/btab727"),
# #   entity = "works",
# #   verbose = TRUE
# # ) %>%
# #   show_works() %>%
# #   knitr::kable()
# #
# #
# #
# nor_bib <- do.call(oa_fetch, bib_ls) %>%
#   oa2bibliometrix()
# nor_bib
#saveRDS(nor_bib, "data/cache_data/nor_bib.RDS")
nor_bib <- readRDS("C:/Users/matthew.grainger/Documents/Projects_in_development/landuse_effects_biodiversity/data/cache_data/nor_bib.RDS")
```


There were `r nrow(nor_bib)` articles included that were conducted in Norway.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "OpenAlex topics weighted by topic score"

topics_aggregated <- map_df(nor_bib$topics, ~ .x %>%
  filter(name == "topic") |> # Filter to include only topics
  select(display_name, score)) |> # Select relevant columns
  group_by(display_name) |> # Group by topic name
  summarise(
    total_score = sum(score), # Sum scores for each topic
    .groups = "drop"
  ) # Drop grouping

# Create a bar plot of the weighted frequencies
topics_aggregated |> 
  top_n(10) |> 
ggplot(aes(x = reorder(display_name, total_score), y = total_score)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Weighted Frequency of Topics by Total Score",
       x = "Topic",
       y = "Total Score (Weighted Frequency)") +
coord_flip()+
  ggthemes::theme_solarized()
```


## Chronological distribution and document contents of included articles

### Geographical distribution of included observations

## Mapping the effects of land use and land cover change on biodiversity

## Mapping the effects of land use and land cover change on ecosystem services

## Mapping the effects of land use and land cover change on carbon storage

## Mapping the effects of land use and land cover change on conflicts and governance

# Limitations of the map

# Limitations of the evidence base

# Conclusions

## Implications for Research

## Implications for management/practice
