---
title: "Effects of land use and land cover changes on biodiversity, ecosystem services, and carbon storage in Norway: A systematic map and decision support tool"
format:
  html:
    toc: true
params:
  version: "0.1"
  search-results: 36418
editor: visual
---

## 

Version: `r params$version`

# Background

Endring i arealbruk og landdekke (LULCC) anses å være den dominerende driveren av nylig tap av biodiversitet både globalt og i Norge, og påvirker direkte naturens evne til å tilby naturgoder og -tjenester (Díaz et al., 2019; Malhi et al., 2020). LULCC påvirker også et økosystems evne til å lagre og binde karbon i jord og biologisk materiale, og påvirker dermed direkte et økosystems evne til å bidra til å motvirke og tilpasse seg klimaendringer. Å kvantifisere, forstå og formidle dynamikken i arealendringer og deres effekter på biodiversitet, økosystemtjenester og karbonlagring er derfor en forutsetning for evidensbaserte arealbruksstrategier. Imidlertid mangler det for øyeblikket systematisk evidens om dette temaet fra norske økosystemer. Mye av den offentlige diskursen på nasjonalt nivå domineres av internasjonale globale kunnskapssammendrag, til tross for betydelig usikkerhet rundt nedskalering og overføring av kunnskap fra globale og regionale synteser til lokalt/nasjonalt nivå, på tvers av økosystemer og taksonomiske grupper. Tidligere kunnskapssammendrag basert på data fra norske økosystemer har enten fulgt spesifikke tilnærminger (f.eks. Økosystemvurdering; (Nybø et al., 2017)), i stor grad basert på ekspertvurderinger (Artsdatabanken, 2021), eller vært begrenset i omfang. Mangelen på en omfattende oversikt over forskningsbasert kunnskap fra norske økosystemer fører også til at ulike interessegrupper og sektorer stoler på ulike "fakta," noe som ofte fører til eskalerende konflikter og dårlige beslutninger. Det er derfor et presserende behov både for en oppsummering av kunnskap og identifisering av reelle kunnskapshull.

Politikk og beslutningstaking knyttet til arealbruk og dens effekter på biodiversitet og klima er for øyeblikket blant de mest konfliktfylte og heftig debatterte politiske spørsmålene i Norge. Konfliktene involverer typisk dilemmaer mellom ulike samfunnsmessige bekymringer og interesser støttet av ulike sektorer og aktørgrupper (Guldbrandsen & Handberg, 2022; Skogen et al., 2021). Kunnskapskontrovers er også ofte en sentral egenskap ved konfliktene, og innebærer uenighet om kunnskapens status og motstridende kunnskapskrav om de miljømessige effektene av ulike alternativer for arealbruk (Aspøy & Stokland, 2022; Rusch et al., 2022). IPBES (2019, 2022a, 2022b) og IPCC (2018) har foreslått styringsmetoder og tiltak, slik som de knyttet til transformasjonsendringer og transformasjonsstyring, for å muliggjøre oppnåelsen av SDG-ene, grensen på 1,5°C og 2050-visjonen for biodiversitet. Slike tilnærminger involverer forbedret integrasjon og samhandling mellom samfunnsmessige mål, styringssektorer og flere kunnskapskilder, og er dermed svært relevante for norsk politikk og beslutningstaking knyttet til arealbruk. Imidlertid hindres utnyttelsen og tilpasningen av internasjonalt anbefalte tilnærminger og verktøy i en norsk kontekst for øyeblikket av mangelen på en omfattende forståelse av nasjonale og lokale dilemmaer og konflikter knyttet til arealbruk, samt de for tiden anvendte verktøyene og tilnærmingene. Det er derfor et presserende behov for en kunnskapssyntese og identifisering av kunnskapshull angående disse aspektene ved arealbruk og dens effekter på biodiversitet og klima.

Systematiske kart (også kjent som Evidenshullkart eller Evidenskart) ble designet for å gi en oversikt over tilgjengelig evidens om et bredt tema (Saran et al., 2020). De søker ikke å svare på smale, veldefinerte spørsmål (slik systematiske oversikter gjør), men heller å gi et middel til å identifisere og katalogisere tilgjengelig evidens om politisk relevante spørsmål (James et al., 2016). Systematiske kart har blitt realisert på en rekke ulike måter, men gir generelt resultatene på en brukervennlig måte, for eksempel ved hjelp av grafer, kart eller søkbare databaser (Miake-Lye et al., 2016). Interaktive plott og kartlegging blir i økende grad brukt for å gi en dynamisk utgang til brukerne (slik som EviAtlas; Haddaway et al., 2019). Det er viktig at systematiske kart ikke fremstilles som et mål i seg selv. Uten ytterligere evidenssyntese må beslutningstakere være forsiktige med å treffe beslutninger basert kun på evidensens vekt alene (dvs. stemmeoptelling, jf. (Hedges & Olkin, 1980)). Systematiske kart kan ikke erstatte rigoriteten til systematiske oversikter, men kan gi beslutningstakere forståelse av evidensgrunnlaget ved å belyse hull eller klynger av evidens. De kan vise hvor evidenssyntese kan finne sted (dvs. det er tilstrekkelig med tilgjengelig evidens for en systematisk oversikt) eller hvor primærforskning er nødvendig (dvs. det er mangel på evidens). Beslutningsmodeller kan supplere resultatene av en systematisk evidenssyntese for å plassere den tilgjengelige evidensen i beslutningskonteksten (Carrick et al., 2019).

## Stakeholder engagement

The outline for the review was presented at a conference held by the Norwegian Environment Agency and the Norwegian Research Council October 23, 2022. A general stakeholder meeting was arranged by NINA on April 25, 2023. The focus and scope of the review was narrowed and better defined during a specific stakeholder event on #date. Details of this meeting and the modifications in scope are available on request. Stakeholders who attended this event were invited to comment on the draft protocol prior to submission for publication. Stakeholders were not engaged during the conduct of the review.

## Objective of the map

The main goal of this project is to provide a comprehensive overview of available knowledge related to the effects of land use and land use changes on Norwegian nature, with a focus on biodiversity, ecosystem services, climate adaptation, carbon sequestration and storage. The map is restricted in geographical scope to Norway, Sweden and Finland (see inclusion criteria below). Moreover, the report is accompanied by a decision support tool as an extension of the systematic map to contextualize the available knowledge. The tool will be designed as a conceptual model of the field and highlight the importance of knowledge gaps in different areas.

### Primary question

The main research question in our systematic map is ‘What evidence exist about the impacts of land use and land cover change on biodiversity, ecosystem services and carbon sequestration in Norway, what type of conflicts have been identified between different forms of land use, and what tools are suggested to improve governance".

### Secondary questions

In addition, we will investigate what characterizes the existing body of literature on the topic, in terms of:

-   What proportion of the research is based on empirical data

-   What proportion of the research is based on experimental vs observational

-   What proportion of the research is based on qualitative vs quantitative/mixed

-   How is research spread across the major ecosystem types/realms

-   How is research spread across the disciplines

-   How is research spread across level of biological organization (species, taxa, ecosystem or services)

-   What are the dominant types of land use that are investigated?

-   What regions (in Norway) are most represented

-   What proportion of the research focuses on governance, conflicts and tools?

-   What are the dominant/marginal social scientific analytical approaches of the studies?

-   What type of actors are most represented in the studies?

-   Which sectors and societal concerns are most studied?

-   Which level of societal scale do the studies represent?

-   Which land use conflicts are most researched?

-   Which tools are proposed and investigated?

# Methods

## Deviations from protocol

## Search for literature

### Online publication databases

Systematic literature searches for academic literature were conducted in Scopus, and in the following Web of Science Core Collection databases:

-   Science Citation Index Expanded (SCI-EXPANDED): 1987–present.

-   Social Sciences Citation Index (SSCI): 1987–present.

-   Arts and Humanities Citation Index (AHCI): 1987–present.

-   Emerging Sources Citation Index (ESCI): 2015–present.

Several preliminary searches with different search terms were performed in Web of Science, which led to adjustments in the search string (See [the protocol](https://osf.io/4ad3x)).

### Search terms and string

We used the PICO framework (James et al., 2016) to structure our search with the search terms and categories derived from the primary research question.

The following search string was used in WoS and a similar string used in Scopus (adjusted to the interface):

<!--# Add link to search strings -->

*TS=("land use" OR "land cover" OR "space use" OR LULCC OR (planning AND (spati\* OR region\* OR area\* OR conservation OR land\* OR urban)) OR (conversion AND (land\* OR habitat\* OR area\* OR ecosystem\* OR natur)) OR (change NEAR/3 (land\* OR area)) OR "national park" OR landscape\* OR habitat\* OR Husbandry OR "land shar" OR "land spar" OR (Outdoor NEAR/3 (life OR livin\* OR recreat)) OR (restoration AND (nature OR ecosystem)) OR forestry OR (reindeer NEAR/3 (herd\* OR husbandr)) OR (livestock NEAR/3 graz) OR "holiday home" OR "secondary home"OR (recreation\* NEAR/3 (home\* OR cottage\* OR cabin\* OR propert)) OR tourism OR "wind power" OR "Land manag" OR "Nature conservation" OR ((governance OR management) AND (spati\* OR region\* OR area\* OR conservation OR land\* OR urban)) OR aquaculture OR "fish farm" OR hydropower OR “green energy” OR “energy development” OR herding OR agriculture OR transportation OR road OR “oil and gas” OR petroleum OR mining OR windpower OR traffic OR road\* OR railway\* OR “hydroelectric power” OR “solar power” OR “solar electricity” OR “power plant” OR LULC OR LULUCF OR AFOLU OR FOLU)*

*AND*

*TS=(sustainab\* OR biodiversity OR "Ecosystem service" OR (carbon NEAR/3 (stor OR sequest)) OR (ecosystem NEAR/3 (impact\* OR effect\* OR provision\* OR regulat\* OR support\* OR cultur)) OR ((loss OR shrink OR bisection\* OR perforation OR fragmentation) AND (habitat OR land)) OR (climate NEAR/3 adapt) OR ecodiversity OR "ecosystem type" OR "ecological process" OR anthropogenic OR conflict OR (diversity NEAR/3 (ecological OR speci\* OR population\*)) OR red-list OR "water purification" OR "soil retention" OR spiritual OR cultural OR aesthetics OR "climate mitigation")*

We screened all systematic reviews and systematic maps from the international literature, the two categories were combined with TS=("Systematic review*" OR "systematic map*" OR "rapid review\*"). For the primary literature, the two categories were combined with TS=(Norway OR Norwegian OR Sweden OR Swedish OR Finland OR Finnish), and all results screened.

### Languages

Studies written in English, Norwegian, Swedish or Danish were included in the review. Studies in other languages were not included in the review.

### Internet searches

### Additional searches

Articles cited by and that cite (forward and backward citation chasing) the 15 benchmark articles were searched for using the citationchaser Shiny App (Haddaway et al. 2022).

### Testing the comprehensiveness of the search

Of the 15 benchmarks, 2 were not recovered in the searches.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Overlap between the different sources of literature and the benchmark papers"

library(CiteSource)
library(tidyverse)
unique_citations <- readRDS("C:/Users/matthew.grainger/Documents/Projects_in_development/landuse_effects_biodiversity/data/unique_citations.RDS")
source_comparison <- compare_sources(unique_citations, comp_type = "sources")
# Generate a source comparison upset plot.
plot_source_overlap_upset(source_comparison, decreasing = c(TRUE, TRUE))
```

### Assembling and managing search results

### Article screening

We used pre-defined inclusion/exclusion criteria to screen eligible studies. Included studies were de-duplicated based on DOI and title matches using the R package 'ASySD' (Hair et al. 2021). We retained the latest version of any version-controlled articles (e.g. published versions of pre-prints were retained).

After de-duplication, title and abstract screening was done using the eligibility criteria, and any studies classified as uncertain were assessed at the full-text stage. The screening was done using Rayyan, with double-screening by a primary and several secondary reviewers. After 50 include/exclude decisions we used the Rayyan-AI 5 star rating to rank the abstracts for assessment. This algorithm assesses abstracts and titles using natural language processing to identify records that have a high potential for inclusion (5 star) or those with a lower potential (1 star).

Weekly monitoring was done for group disagreement rates, and the eligibility criteria were re-evaluated and readjusted when the rate exceeded 15%. Conflicting decisions were reconciled through discussions with the whole review group. Studies tagged as unsure at each stage were assessed at the next level, with those remaining tagged as unsure reviewed by the entire team. 

### Eligibility criteria

We reviewed the collated studies obtained from the searches using the criteria set below. Only studies conducted in Norway, Sweden and Finland or systematic reviews (also referred to as meta-analyses) with global, European or Fennoscandian coverage were included in the review. Only records where the full-text is written in English, Norwegian, Swedish, or Danish were considered eligible.

Studies will be eligible if they assess land use or land-use-change effects on the biodiversity, ecosystem services/functions or carbon sequestration and storage in Feno-Scandinavia. Only studies that focus on terrestrial, aquatic or coastal areas are include (i.e. studies focusing on marine systems are excluded).

Studies will also be eligible if they assess conflicts, tools or governance relating to land use and cover change its effects on biodiversity, ecosystem services and/or carbon sequestration and storage by ecosystems in Fenno-Scandinavia. The relation should be explicit in study for the record to be included.

Studies of all scales (national, regional, and local) will be considered eligible.

2.  **Eligible interventions**

Land use and/or land cover change through one or multiple drivers (see Appendix) which are anthropogenic in origin and result in one or more of the core mechanisms (change in habitat – amount; habitat composition; habitat structure; habitat – function; indirect, e.g. changed human use in recreation etc).

3.  **Eligible outcomes**

Eligible outcomes will include function, structure, diversity, abundance, behaviour, composition, or richness in any level of biological organisation (genetic, species, community, ecosystem, landscape – multiple ecosystems), as well as studies focusing on the amount of, diversity or category of ecosystem services (provisioning, regulating, supporting or cultural service and, carbon storage and sequestration).

In addition, studies within the social sciences will be included if they explicitly address conflicts, tools and governance pertaining to effects land use and land cover change on biodiversity, ecosystem services and/or carbon sequestration and storage.

4.  **Eligible types of study design**

All study designs will be considered eligible, including both quantitative and qualitative studies.

### Study validity assessment

### Data coding strategy

### Data mapping method

# Review findings

## Results of the search for literature

```{r, Prisma statement}
#| echo: false
#| message: false
#| warning: false
library(PRISMAstatement)

prisma(
  found = params$`search-results`,
  found_other = 1965,
  no_dupes = 20265,
  screened = 14808, # Needs updating
  screen_exclusions = 12144,
  full_text = 2398,
  full_text_exclusions = 0,
  qualitative = 0,
  quantitative = 0,
  width = 800, height = 800
)
```

### Sources and document types of included articles

## Review descriptive statistics

## Norway papers

```{r}
#| echo: false
#| message: false
#| warning: false
# Extract data and plot from the extraction sheet ----
## load libraies----
library(tidyverse, quietly = TRUE)
library(googlesheets4, quietly = TRUE)
library(rayyanR)
library(synthesisr)
library(openalex)
library(openalexR)
# gs4_auth()
## read in GoogleSheets----

form <- read_sheet("https://docs.google.com/spreadsheets/d/1wwOfwz_QqwXpNxKx4Jupn336x2GttOfoXKibd-_rMuc")

## clean names----
form <- form |>
  janitor::clean_names()
## explore----
#names(form)

# ### select Norwegian studies----
Nor_form <- form |>
  filter(grepl("Norway", country))

### selet studies that are included in the review----
Nor_form_inc <- Nor_form |>
  filter(does_the_study_assess_land_use_or_land_use_change_effects_on_the_biodiversity_ecosystem_services_functions_or_carbon_sequestration_and_storage_in_terrestrial_aquatic_or_coastal_areas_are_include_i_e_studies_focusing_on_marine_systems_are_excluded == "Yes" | does_the_study_assess_conflicts_tools_or_governance_relating_to_land_use_and_cover_change_its_effects_on_biodiversity_ecosystem_services_and_or_carbon_sequestration_and_storage_by_ecosystems_in_fenno_scandinavia_the_relation_should_be_explicit_in_study_for_the_record_to_be_included == "Yes")

### Get matched details from Rayyan----

# read the Rayyan results

rayyan_biblio <- synthesisr::read_ref("data/Extraction/articles.ris")
rayyan_biblio_cleaned <- rayyanR::parse_rayyan(rayyan_df = rayyan_biblio)

Nor_form_inc$accession_zr<-unlist(Nor_form_inc$study_id)

join_Nor_form <- Nor_form_inc |>
  left_join(rayyan_biblio_cleaned, join_by("accession_zr"))

# join_Nor_form |> view()

join_Nor_form |>
  filter(is.na(join_Nor_form$doi)) |>
  select(study_id, study_title)

#### find missing doi's by title
which(is.na(join_Nor_form$doi))
#11, 35, 47,52,60,61,63

works_search <- oa_fetch(
  entity = "works",
  title.search = c("Roads forestry wolves"),
  verbose = TRUE
)
works_search$doi
# "https://doi.org/10.1002/ecs2.3358"


works_search <- oa_fetch(
  entity = "works",
  title.search = c("CONFLICT AND COMPLEXITY - THE CASE OF HYDROPOWER CONSTRUCTION IN NORWAY"),
  verbose = TRUE
)
works_search$doi
# "https://doi.org/10.3233/hsm-1986-6105"
works_search <- oa_fetch(
  entity = "works",
  title.search = c("Influence of site and stand factors on the distribution of crustose lichens of the Caliciales in a suboceanic spruce forest area in central Norway"),
  verbose = TRUE
)
works_search |> view()
works_search$doi
# "https://doi.org/10.1017/s0024282996000424"

works_search <- oa_fetch(
  entity = "works",
  title.search = c("Lichens in spruce forest stands of different successional stages in central Norway with emphasis on diversity and old growth species"),
  verbose = TRUE
)

works_search$doi
works_search <- oa_fetch(
  entity = "works",
  title.search = c("Impacts of experimental habitat fragmentation on ground beetles"),
  verbose = TRUE
)

works_search


works_search <- oa_fetch(
  entity = "works",
  title.search = c("Cumulative impacts of tourist resorts on wild reindeer"),
  verbose = TRUE
)

works_search$doi

#Home range size and choice of management strategy for lynx in Scandinavia

works_search <- oa_fetch(
  entity = "works",
  title.search = c("Home range size and choice of management strategy for lynx in Scandinavia"),
  verbose = TRUE
)

works_search$doi
join_Nor_form[11, ]$doi <- "10.3233/hsm-1986-6105"
join_Nor_form[35, ]$doi <- "10.1002/ecs2.3358"
join_Nor_form[47, ]$doi <- "10.1017/s0024282996000424"
join_Nor_form[52, ]$doi <- "10.1127/nova.hedwigia/66/1998/283"
join_Nor_form[60, ]$doi <- NA
join_Nor_form[61, ]$doi <- "10.14430/arctic829"
join_Nor_form[61, ]$doi<-"10.1007/s002670010195"
# get openAlexR data


doi_list <- strsplit(join_Nor_form$doi, split = " ")
doi_list <- sapply(doi_list, "[[", 1)

# # oa_fetch(
# #   doi = c(doi_list),
# #   entity = "works",
# #   verbose = TRUE
# # ) %>%
# #   show_works() %>%
# #   knitr::kable()

bib_ls <- list(
  doi = c(doi_list),
  entity = "works"
)

# # oa_fetch(
# #   # identifier = c("W2741809807", "W2755950973"),
# #   doi = c("10.1016/j.joi.2017.08.007", "https://doi.org/10.1093/bioinformatics/btab727"),
# #   entity = "works",
# #   verbose = TRUE
# # ) %>%
# #   show_works() %>%
# #   knitr::kable()
# #
# #
# #
nor_bib <- do.call(oa_fetch, bib_ls) %>%
  oa2bibliometrix()
nor_bib
saveRDS(nor_bib, "data/cache_data/nor_bib.RDS")
saveRDS(join_Nor_form, "data/cache_data/join_Nor_form.RDS")
nor_bib <- readRDS("C:/Users/matthew.grainger/Documents/Projects_in_development/landuse_effects_biodiversity/data/cache_data/nor_bib.RDS")
join_Nor_form<-readRDS("C:/Users/matthew.grainger/Documents/Projects_in_development/landuse_effects_biodiversity/data/cache_data/join_Nor_form.RDS")
```


There were `r nrow(nor_bib)` articles included that were conducted in Norway.

```{r openAlex topics/concepts}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "OpenAlex topics weighted by topic score"

topics_aggregated <- map_df(nor_bib$topics, ~ .x %>%
  filter(name == "topic") |> # Filter to include only topics
  select(display_name, score)) |> # Select relevant columns
  group_by(display_name) |> # Group by topic name
  summarise(
    total_score = sum(score), # Sum scores for each topic
    .groups = "drop"
  ) # Drop grouping

# Create a bar plot of the weighted frequencies
topics_aggregated |> 
  top_n(10) |> 
ggplot(aes(x = reorder(display_name, total_score), y = total_score)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Weighted Frequency of Topics by Total Score",
       x = "Topic",
       y = "Total Score (Weighted Frequency)") +
coord_flip()+
  ggthemes::theme_solarized()
```

## Chronological distribution and document contents of included articles

```{r publication year}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Cumulative sum of Norwegian publications per year"

nor_bib |> 
  mutate(year=year(publication_date)) |> 
  group_by(year) |> 
  tally() |> 
  mutate(csum=cumsum(n)) |> 
  ggplot(aes(year,csum)) +
  geom_point(colour="blue")+
  labs(x="Publication Year", y="Cumulative sum of publications")+
  ggthemes::theme_solarized()
```

### Geographical distribution of included observations


```{r County distribution}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Norwegian counties where research was undertaken"
my_list<-strsplit(join_Nor_form$county_name_of_norwegian_county, ",")

# Function to replace specific problematic characters
replace_problematic_chars <- function(x) {
  # Replacing specific problematic unicode sequence if visible in the string
  x <- gsub("\xc3\x98", "O", x, fixed = TRUE)  # Replacing 'Ø' which sometimes appears as '\xc3\x98' in UTF-8

  return(x)
}

# Apply this function to each element in the list
my_list <- lapply(my_list, function(x) {
  if (is.character(x)) {
    sapply(x, replace_problematic_chars, USE.NAMES = FALSE)
  } else {
    x
  }
})

# Flatten the list
flattened <- unlist(my_list)

# Remove NA and clean up
flattened <- na.omit(flattened)
flattened <- trimws(flattened)

# Create a dataframe
df <- data.frame(Name = flattened, stringsAsFactors = FALSE)

# Print the dataframe
#print(df)

df |> 
  group_by(Name) |> 
  tally() |> 
  ggplot(aes(reorder(Name, n),n)) + 
  geom_histogram(stat="identity", fill="blue")+
  coord_flip()+
  labs(x="", y="Antall publikasjoner")+
  ggthemes::theme_solarized()

```


```{r topic model}
#| echo: false
#| message: false
#| warning: false

library(ldatuning)
library(topicmodels)
library(tm)
library(tidytext)

jn_dtm<-join_Nor_form |> 
  select(study_id, title, abstract) |> 
  mutate(title_abs=paste0(title, abstract))


jn_dtm$study_id<-unlist(jn_dtm$study_id)

jn_dtm<-jn_dtm |> 
  unnest_tokens(word,title_abs)

data(stop_words)

jn_dtm <- jn_dtm %>%
  anti_join(stop_words)


jn_dtm

```























## Mapping the effects of land use and land cover change on biodiversity, carbon and conflicts

```{r keyword co-occurrence}
library(bibliometrix)
# Create keyword co-occurrences network

NetMatrix <- biblioNetwork(nor_bib, analysis = "co-occurrences", network = "keywords", sep = ";")

# Plot the network
net=networkPlot(NetMatrix, normalize="association", weighted=T, n = 30, Title = "Keyword Co-occurrences", type = "fruchterman", size=T,edgesize = 5,labelsize=0.7)
```



```{r topic modelling}

```



#

# Limitations of the map

# Limitations of the evidence base

# Conclusions

## Implications for Research

## Implications for management/practice
